### 2019-02-26 rick/1KGP.Rmd
update

---
title: "HGP analysis"
author: 
- name: Rick Farouni
  affiliation:
  - &cruk Department of Human Genetics, McGill University,  Montreal, Canada
date: '`r format(Sys.Date(), "%Y-%B-%d")`'
output:
  html_notebook:
    df_print: paged
    code_folding: show
    toc: yes
    toc_float: 
      collapsed: false
      smooth_scroll: false
---

# Preparations

```{r setup}
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file(), fig.width=16)
options(scipen=999)
```

```{r}
library(Matrix)
library(irlba)
library(threejs)
library(tidyverse)
library(lfa)
```

# Download data

Download a fast parser program (after compilation you'll have a program called a.out)

```{bash}
wget https://raw.githubusercontent.com/bwlewis/1000_genomes_examples/master/parse.c
cc -O2 parse.c
```

Download 1000 genomes phenotype data file

```{bash}
wget ftp://ftp.1000genomes.ebi.ac.uk/vol1/ftp/technical/working/20130606_sample_info/20130606_g1k.ped
```

Download 1000 genomes genotype data file for chr 20

```{bash}
wget ftp://ftp-trace.ncbi.nih.gov/1000genomes/ftp/release/20130502/ALL.chr20.phase3_shapeit2_mvncall_integrated_v5a.20130502.genotypes.vcf.gz

```

Read just the header of the chromosome file to obtain the sample identifiers

```{r}
ids <- readLines(pipe("zcat ALL.chr20.phase3_shapeit2_mvncall_integrated_v5a.20130502.genotypes.vcf.gz  |
                      sed -n /^#CHROM/p | 
                      tr '\t' '\n' | 
                      tail -n +10"))
head(ids)
```

Write ids to file

```{r}
writeLines(ids, "ids.txt")
```

Download bas files for all the samples

```{bash message=FALSE}
for id in `cat ids.txt`
  do
  wget ftp://ftp.1000genomes.ebi.ac.uk/vol1/ftp/phase3/data/${id}/alignment/${id}.chrom20.*.bam.bas
done
wait
```

# Read bas files

```{r}
read_basfiles <- function(bas_filepath) {

  bas_file <- read_tsv(bas_filepath,
                  col_types = cols_only('sample' = 'c', 
                                        'platform' = 'c',        
                                       'average_quality_of_mapped_bases' = col_guess())) 
  return( bas_file)
}
```

```{r}
bas_files <- list.files(path = ".",
                        pattern="*bam.bas")

bas_dt <- 
  map_dfr(bas_files,read_basfiles ) %>%
  rename(avg_qual="average_quality_of_mapped_bases")
bas_dt
```

```{r}
bas_dt <-
  bas_dt %>%
  group_by(sample) %>%
  mutate(id=1:n()) %>%
  select(-platform)%>%
  spread(id,avg_qual)
bas_dt
```
```{r}
n_not_na <- function(x) {
 sum(!is.na(x))
}
```

```{r}
 bas_dt <- 
  bas_dt %>%
  ungroup()%>% 
    mutate( avg_qual_sd =pmap_dbl(select(., -sample), lift_vd(sd,na.rm =TRUE )),
           avg_qual_mean = rowMeans(select(., -sample),na.rm =TRUE),
           avg_qual_n= pmap_dbl(select(., -sample), lift_vd(n_not_na)),
           avg_qual_min= pmap_dbl(select(., -sample), lift_vd(min, na.rm =TRUE)),
           avg_qual_max= pmap_dbl(select(., -sample), lift_vd(max, na.rm =TRUE))) %>%
  select (sample, avg_qual_n, avg_qual_min,avg_qual_max,avg_qual_mean,avg_qual_sd, everything() ) 
```

```{r}
ped <- 
  read_tsv(url("ftp://ftp.1000genomes.ebi.ac.uk/vol1/ftp/technical/working/20130606_sample_info/20130606_g1k.ped"))%>% 
  rename(sample="Individual ID")
ped 
```

```{r}
bas_dt <- inner_join( ped  %>% 
                       select(sample, Population),
                       bas_dt,
                     by="sample")
bas_dt
```

```{r}
write_tsv(bas_dt, "average_quality_chrom21.txt")
```

## Plot

```{r fig.width=20}
ggplot(bas_dt,
       aes(x=avg_qual_min,
           y=avg_qual_max, 
           colour=avg_qual_n )) +
  geom_point(size=2)

```

```{r fig.width=20}
ggplot(bas_dt,
       aes(x=avg_qual_mean,
           y=avg_qual_sd, 
           colour=avg_qual_n )) +
  geom_point(size=2)

```

# Load Gentype data

Load VCF file into a sparse matrix

```{r}
p <- pipe("zcat ALL.chr20.phase3_shapeit2_mvncall_integrated_v5a.20130502.genotypes.vcf.gz  |
          sed /^#/d  |
          cut  -f '10-' |
          ./a.out | 
          cut -f '1-2'")
x <- read.table(p,
                colClasses=c("integer","integer"), 
                fill=TRUE,
                row.names=NULL)
```

```{r}
# Convert to a sparse matrix of people (rows) x variant (columns)
chr20 <- sparseMatrix(i=x[,2], j=x[,1], x=1.0)

# Inspect the dimensions of this matrix
print(dim(chr20))
```

The matrix has 2,504 rows (people) and 1,812,841 columns (variants)

# PCA

compute the first three principal component vectors

```{r}
cm <- colMeans(chr20)
p <- irlba(chr20,
           nv=3,
           nu=3,
           tol=0.1,
           center=cm)
```

 add to dataset
```{r}
colnames(p$u) <- paste0("pc",1:3)
bas_dt <- bind_cols(bas_dt, 
                    as_tibble(p$u))
```

```{r}
ggplot(bas_dt,
       aes(pc1, pc3, color=Population)) +
  geom_point() +
  theme_bw() 
```

# Logistic factor analysis

```{r}
chr20_mtx <- t(as.matrix(chr20[,1:50000]))
```

compute the first three factors including the intercept

```{r}
LF <- lfa(chr20_mtx, 4)
dim(LF)
```

```{r}

bas_dt <- bind_cols(bas_dt,
                    tibble(lf1=LF[,1], lf2=LF[,2],lf3=LF[,3]))
```

```{r}
ggplot(bas_dt,
       aes(lf1, lf2, color=Population)) +
  geom_point() +
  theme_bw()
```

computes the individual-specific allele frequencies

```{r}
subset <- af(chr20_mtx, LF)

subset[1:5,1]
```

```{r}
sessionInfo()
```