### 2019-02-27 rick/1kGP_lukeMod.Rmd
Added a modified version of the R script from Rick

---
title: "Individual Specific Allele Frequency"
author: "Rick + Luke"
date: "2/27/2019"
output: html_document
---

# Preparations

```{r}
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file(), fig.width=16)
options(scipen=999)
```
```{bash}
wget https://raw.githubusercontent.com/bwlewis/1000_genomes_examples/master/parse.c
cc -O2 parse.c
```

```{r}
library(Matrix)
library(irlba)
library(threejs)
library(tidyverse)
library(lfa)

library(glm2)
```

Read just the header of the chromosome file to obtain the sample identifiers

```{r}
GenoPath='/Users/luke/genomes/genomes/hg19/phase3/'

ids <- readLines(pipe(paste("gzcat ",GenoPath,"ALL.chr22.phase3_shapeit2_mvncall_integrated_v5a.20130502.genotypes.vcf.gz  |
                      sed -n /^#CHROM/p | 
                      tr '\t' '\n' | 
                      tail -n +10",sep='')))
head(ids)
```

Write ids to file

```{r}
Dir = '/Users/luke/Documents/QualityPaper/rick/'
setwd(Dir)
writeLines(ids, "ids.txt")
```

# Read bas files

```{r}
read_basfiles <- function(bas_filepath) {

  bas_file <- read_tsv(bas_filepath,
                  col_types = cols_only('sample' = 'c',        
                                       'average_quality_of_mapped_bases' = col_guess())) 
  return( bas_file)
}
```

```{r}
bas_files <- list.files(path = "/Users/luke/genomes/bas_file/1000GenomesBAS/",
                        pattern="*bam.bas", full.names = TRUE)

bas_dt <- 
  map_dfr(bas_files,read_basfiles ) %>%
  rename(avg_qual="average_quality_of_mapped_bases") %>%
  group_by(sample) %>% summarize(avg_qual = mean(avg_qual))

#keep samples in 1kGP VCF files
bas_dt <- bas_dt[which((bas_dt$sample %in% ids)==T),]

bas_dt
```
Get Sample Info, 
```{r}
ped <- 
  read_tsv(url("ftp://ftp.1000genomes.ebi.ac.uk/vol1/ftp/technical/working/20130606_sample_info/20130606_g1k.ped"))%>% 
  rename(sample="Individual ID")

bas_dt <- inner_join( ped  %>% 
                       select(sample, Population),
                       bas_dt,
                     by="sample")
bas_dt
```

# Load Gentype data

Load VCF file into a sparse matrix

```{r}
p <- pipe(paste("gzcat ",GenoPath,"ALL.chr22.phase3_shapeit2_mvncall_integrated_v5a.20130502.genotypes.vcf.gz  |
          sed /^#/d  |
          cut  -f '10-' |
          ./a.out | 
          cut -f '1-2'",sep=''))
x <- read.table(p,
                colClasses=c("integer","integer"), 
                fill=TRUE,
                row.names=NULL)
```

```{r}
# Convert to a sparse matrix of people (rows) x variant (columns)
chr22 <- sparseMatrix(i=x[,2], j=x[,1], x=1.0)

# Inspect the dimensions of this matrix
print(dim(chr22))
```

# PCA

compute the first three principal component vectors

```{r}
cm <- colMeans(chr22)
p <- irlba(chr22,
           nv=3,
           nu=3,
           tol=0.1,
           center=cm)
```

 add to dataset
```{r}
colnames(p$u) <- paste0("pc",1:3)
bas_dt <- bind_cols(bas_dt, 
                    as_tibble(p$u))
```

```{r}
ggplot(bas_dt,
       aes(pc1, pc3, color=Population)) +
  geom_point() +
  theme_bw() 
```

# Logistic factor analysis

```{r}
chr22_mtx <- t(as.matrix(chr22[,1:50000]))
```

compute the first three factors including the intercept

```{r}
LF <- lfa(chr22_mtx, 4)
dim(LF)
```

```{r}

bas_dt <- bind_cols(bas_dt,
                    tibble(lf1=LF[,1], lf2=LF[,2],lf3=LF[,3]))
```

computes the individual-specific allele frequencies

```{r}
subset <- af(chr22_mtx, LF)

subset[1:5,1]
```

```{r}
ggplot(bas_dt,
       aes(lf1, lf2, color=Population)) +
  geom_point() +
  theme_bw()
```

Logistic Regression
```{r}
LF = mapply(function(x, y){
                           glm(x 
                                ~ bas_dt$avg_qual 
                                + offset(y))}, 
                           as.data.frame(t(chr22_mtx)),
                           as.data.frame(t(subset)))

```

```{r}
Pop = apply(head(chr22_mtx,1000), 1, function(x)
                    glm2(x 
                          ~ bas_dt$Population
                          + bas_dt$avg_qual))
```