### 2019-03-11 rick/final/5_make_plots.Rmd

---
title: "HGP analysis: plot"
author: 
- name: Rick Farouni, 
  affiliation:
  - &cruk Department of Human Genetics, McGill University,  Montreal, Canada
date: '`r format(Sys.Date(), "%Y-%B-%d")`'
output:
  html_notebook:
    df_print: paged
    code_folding: show
    toc: yes
    toc_float: 
      collapsed: false
      smooth_scroll: false
---

# Preparations

```{r setup}
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file(), fig.width=16)
options(scipen=999)
```

```{r}
library(tidyverse)
library(lfa)
library(glm2)
library(furrr)
library(broom)
library(tictoc)
options(future.globals.maxSize= 1000000000000)
plan(multicore)
```

```{r}
gc()
```

```{r}
bas_filepath <- "./data/metadata/average_quality_dt.txt"
bas_dt <- read_tsv(bas_filepath)
avg_qual_mean <- bas_dt$avg_qual_mean
```
```{r}
head(avg_qual_mean)
```

```{r}
S <-30
m <- 76871
subset <- vector("list", length = S)
for(i in 1:S){
  subset[[i]] <-readRDS(file.path("./data/preprocessed/subsets", paste0("fits_subset_", i, ".rds")))
}
```

```{r}
gc()
```

```{r}

get_plot <- function(data) {
  p <- ggplot(bind_cols(data,
                        bas_dt %>%
                          select(Population,
                                 avg_qual_mean))) +
    geom_point(aes(avg_qual_mean,
                   genotype),
               size = .5,
               alpha = 0.4) +
    geom_point(aes(avg_qual_mean,
                  iaf,
                  colour = "af")) +
    geom_line(aes(avg_qual_mean,
              p_0,
              colour = "p_0"),
          linetype = "dashed") +
    geom_line(aes(avg_qual_mean,
                  p_1,
                  colour = "p_1")) +
    theme_bw() +
    facet_wrap( ~ Population)

  return(p)
}

data_list_subset <- 
  data_list_sub %>%
    mutate(plotted= future_map(data, get_plot)) 
data_list_subset
```

```{r, fig.width=20}
data_list_subset %>%
  arrange(dev_logp )%>%
  slice(1:5)%>%
  pull(plotted)
```

```{r}
sessionInfo()
```