### 2019-05-01 rick/misc/1KGP_v2.Rmd
minor text and major clean up of git repo

---
title: "HGP analysis: fit model"
author: 
- name: Rick Farouni, 
  affiliation:
  - &cruk Department of Human Genetics, McGill University,  Montreal, Canada
date: '`r format(Sys.Date(), "%Y-%B-%d")`'
output:
  html_notebook:
    df_print: paged
    code_folding: show
    toc: yes
    toc_float: 
      collapsed: false
      smooth_scroll: false
---

# Preparations

```{r setup}
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file(), fig.width=16)
options(scipen=999)
```

```{r}
library(tidyverse)
library(lfa)
library(furrr)
library(broom)
library(tictoc)
options(future.globals.maxSize= 1000000000000)
plan(multicore)
```

```{r}
gc()
```

```{r}
gcinfo(TRUE) 
```

#  Fit model

```{r}

to_datalist <- function(i){

    data <- tibble(genotype=data_genotype[i,],
               iaf=data_iaf[i,])

     return(data)       
}

get_predicted_mean <- function(model_fit) {
 predicted <- 
   augment(model_fit, type.predict = "response") %>%
   select("genotype","offset.iaf.",".fitted") %>%
   rename(iaf="offset.iaf.", mu=".fitted" )
  return(predicted)
}

get_stats <- function(model_fit){
  stats_dt <- 
    tidy(model_fit)  %>%
    filter(term=="avg_qual_mean") %>%
    select(estimate, p.value) %>%
    mutate(deviance= anova(model_fit)["Deviance"][[1]][2])
  return(stats_dt)
}

fit_model <- function(dataset){

   fit <- glm(genotype ~ avg_qual_mean + offset(iaf),
                           family ="binomial",
                           data=dataset)

   stats <- get_stats(fit)
   predicted <- get_predicted_mean(fit)

   return(list(stats, predicted))
}

get_plot <- function(data){

 p <- ggplot(bind_cols(data, 
                   bas_dt %>% 
                     select(Population,
                            avg_qual_mean))) +
    geom_point(aes(avg_qual_mean, 
                   genotype),
               size=.5,
               alpha= 0.4) +
    geom_line(aes(avg_qual_mean,
                  iaf, 
                  colour="af"),
              linetype="dashed") +
    geom_line(aes(avg_qual_mean,
                  mu,
                  colour="mu")) +
      theme_bw() +
    facet_wrap(~Population)

  return(p)
}
```

```{r}
bas_filepath <- "./data/metadata/average_quality_dt.txt"
bas_dt <- read_tsv(bas_filepath)
avg_qual_mean <- bas_dt$avg_qual_mean
```

```{r}
fit_data <- function(data_list){

  message("fitting model")
  tic()
  data_list <- 
    data_list %>%
  #slice(1:10) %>%
  transmute(
    locus=locus,
    fit1 = future_map(dataset, fit_model))  %>%
  transmute(
    locus=locus,
    stats = future_map(fit1, c(1)),
    predicted =future_map(fit1, c(2))) %>%
    unnest(stats) %>%
    arrange(-deviance)

  toc()

  return(data_list)

}

```

```{r}
  data_genotype <- read.bed("./data/plink/ALL_Pops")
  message("plink data loaded")

  # compute the first d factors including the intercept

  data_iaf <- readRDS("./data/preprocessed/iaf.rds")

  loci <- c(1:dim(data_genotype)[1])
  names(loci) <- loci

  message("to datalist")

  tic()
  data_list <- future_map(loci, 
                 to_datalist)

  toc()

  data_list <- 
  enframe(data_list,
          name = "locus", 
          value = "dataset")

saveRDS(data_list, "./data/preprocessed/data_list.rds")
```

```{r}
data_list <- fit_data (data_list)
data_list
```

```{r save_data}
saveRDS(data_list, "./data/preprocessed/data_fits.rds")
```

```{r}
n <-20
target_loci <- 
  data_list %>% 
  slice(1:n) %>% 
  pull(locus)
target_loci 
```

```{r}
data_list_subset <- 
  data_list %>%
   filter(locus %in% target_loci ) %>%
    mutate(plotted= future_map(predicted, get_plot)) 
data_list_subset
```

```{r}
data_list_subset %>%
  pull(plotted)
```

```{r}
sessionInfo()
```