### 2019-05-01 RunAnalysis/2_run_lfa.Rmd
more git cleaning

---
title: "HGP analysis: LFA"
author: 
- name: Rick Farouni, 
  affiliation:
  - &cruk Department of Human Genetics, McGill University,  Montreal, Canada
date: '`r format(Sys.Date(), "%Y-%B-%d")`'
output:
  html_notebook:
    df_print: paged
    code_folding: show
    toc: yes
    toc_float: 
      collapsed: false
      smooth_scroll: false
---

# Preparations

```{r setup}
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file(), fig.width=16)
options(scipen=999)
```

```{r}
library(lfa)
library(tictoc)
```
```{r}
gc() #- do it now
gcinfo(TRUE) #-- in the future, show when R does it
```

```{bash}
free -t -g
```

```{r load}
tic()
 data <- read.bed("./data/plink/ALL_Pops")
toc()
 message("plink data loaded")
```

```{bash}
free -t -g
```

```{r lfa}
tic()
LF <- lfa(data, 5)
toc()
message("lfa run complete")
```

```{r save}
tic()
saveRDS(LF, "./data/preprocessed/lf5.rds")
toc()
```
```{r}
gcinfo(verbose = FALSE) #-- don't show it anymore
gc()
```

```{r}
sessionInfo()
```